---
title: "Seroprevalence and Vaccination Coverage"
author: "Stuart Torres, Ben Arnold"
date: "Updated `r Sys.Date()`"
project:
  output-dir: output
format:
  html:
    toc: true
    embed-resources: true
    code-fold: true
    link-external-newwindow: true
---

# Preamble

```{r preamble, message=FALSE}
library(here)
source(here("R", "0-config.R"))
```

# Load and prepare data

```{r load-data}
#-------------------------------
# read processed dataset
#-------------------------------
d0 <- read_rds(here("data", "ecomid_sero_vpd_2026-06-11.rds"))

#-------------------------------
# collapse 4-level community_type to 3 levels
# Esmeraldas City | Borbon | Rural Villages
# using community_name as the source of truth
#-------------------------------
d0 <- d0 %>%
  mutate(
    community_type_3 = case_when(
      community_name == "Esmeraldas" ~ "Esmeraldas City",
      community_name == "Borbon"     ~ "Borbon",
      TRUE                           ~ "Rural Villages"
    ),
    community_type_3 = factor(community_type_3,
                               levels = c("Esmeraldas City", "Borbon", "Rural Villages")),
    pathogen = factor(pathogen,
                      levels = c("Tetanus", "Diphtheria", "Rubella", "Measles")),
    timepoint = as.integer(timepoint),
    timept_f  = factor(timepoint, levels = c(6, 9, 12, 18, 24))
  )

cat("Total rows:", nrow(d0), "\n")
cat("Unique children:", n_distinct(d0$childid), "\n")
cat("Unique child-visits (per pathogen):", n_distinct(d0$childid, d0$timepoint), "\n")
```

```{r helpers}
#-------------------------------
# helper: compute exact binomial CI and format cell text
# used in all seroprevalence and coverage tables (7+ tables)
#-------------------------------
add_binom_ci <- function(df) {
  df %>%
    rowwise() %>%
    mutate(
      prev_pct = if_else(n_test > 0, n_pos / n_test * 100, NA_real_),
      ci_lo    = if_else(n_test > 0, binom.test(n_pos, n_test)$conf.int[1] * 100, NA_real_),
      ci_hi    = if_else(n_test > 0, binom.test(n_pos, n_test)$conf.int[2] * 100, NA_real_),
      cell     = if_else(
        n_test > 0,
        sprintf("%d/%d (%.1f%%) [%.1f, %.1f]", n_pos, n_test, prev_pct, ci_lo, ci_hi),
        "0/0 (—)"
      )
    ) %>%
    ungroup()
}

#-------------------------------
# helper: compute seroprevalence cells for one pathogen
# returns long-format with timepoint_lbl and community_type_3
#-------------------------------
seroprev_cells <- function(data, pathogen_name) {
  dp <- data %>% filter(pathogen == pathogen_name)

  # by community × timepoint
  by_ct <- dp %>%
    group_by(community_type_3, timepoint) %>%
    summarise(n_pos = sum(seropos, na.rm = TRUE),
              n_test = sum(!is.na(seropos)), .groups = "drop")

  # overall column: pool across communities, by timepoint
  by_t <- dp %>%
    group_by(timepoint) %>%
    summarise(n_pos = sum(seropos, na.rm = TRUE),
              n_test = sum(!is.na(seropos)), .groups = "drop") %>%
    mutate(community_type_3 = factor("Overall"))

  # overall row: pool across timepoints, by community
  by_c <- dp %>%
    group_by(community_type_3) %>%
    summarise(n_pos = sum(seropos, na.rm = TRUE),
              n_test = sum(!is.na(seropos)), .groups = "drop") %>%
    mutate(timepoint = 0L)

  # grand total
  grand <- dp %>%
    summarise(n_pos = sum(seropos, na.rm = TRUE),
              n_test = sum(!is.na(seropos))) %>%
    mutate(community_type_3 = factor("Overall"), timepoint = 0L)

  bind_rows(by_ct, by_t, by_c, grand) %>%
    add_binom_ci() %>%
    mutate(timepoint_lbl = if_else(timepoint == 0L, "Overall",
                                   paste0(timepoint, " months")))
}

#-------------------------------
# helper: compute vaccine coverage cells for one dose indicator
# vacc_var: character name of visit-specific vaccination variable (e.g. "penta1_visit")
# dose_label: label for this dose (e.g. "Penta 1")
#-------------------------------
coverage_cells <- function(data, vacc_var, dose_label) {
  dc <- data %>%
    mutate(
      vacc = if_else(.data[[vacc_var]] == "YES", 1L, 0L)
    )

  by_ct <- dc %>%
    group_by(community_type_3, timepoint) %>%
    summarise(n_pos = sum(vacc, na.rm = TRUE),
              n_test = sum(!is.na(vacc)), .groups = "drop")

  by_t <- dc %>%
    group_by(timepoint) %>%
    summarise(n_pos = sum(vacc, na.rm = TRUE),
              n_test = sum(!is.na(vacc)), .groups = "drop") %>%
    mutate(community_type_3 = factor("Overall"))

  by_c <- dc %>%
    group_by(community_type_3) %>%
    summarise(n_pos = sum(vacc, na.rm = TRUE),
              n_test = sum(!is.na(vacc)), .groups = "drop") %>%
    mutate(timepoint = 0L)

  grand <- dc %>%
    summarise(n_pos = sum(vacc, na.rm = TRUE),
              n_test = sum(!is.na(vacc))) %>%
    mutate(community_type_3 = factor("Overall"), timepoint = 0L)

  bind_rows(by_ct, by_t, by_c, grand) %>%
    add_binom_ci() %>%
    mutate(timepoint_lbl = if_else(timepoint == 0L, "Overall",
                                   paste0(timepoint, " months")),
           dose = dose_label)
}

#-------------------------------
# helper: pivot long cells to display-ready wide table
# id_col: row label column
# name_col: column to pivot on
# value_col: cell text column
# col_order: desired column order after pivot
#-------------------------------
pivot_wide_table <- function(df, id_col, name_col, value_col, col_order) {
  df %>%
    select(all_of(c(id_col, name_col, value_col))) %>%
    pivot_wider(names_from = all_of(name_col), values_from = all_of(value_col)) %>%
    select(all_of(id_col), any_of(col_order))
}

comm_col_order <- c("Esmeraldas City", "Borbon", "Rural Villages", "Overall")
```

---

# Table 1a: Household and participant characteristics

```{r table-1a}
#-------------------------------
# load survey data and join to
# child-level community info from serology
#-------------------------------
survey <- read_csv(here("data", "surveydata_20251003.csv"), show_col_types = FALSE)

child_comm <- d0 %>%
  distinct(childid, house_ID, community_type_3)

hh <- child_comm %>% 
  left_join(survey, by = "house_ID")

cat("Children in serology data:", n_distinct(child_comm$childid), "\n")
cat("Children matched to survey:", sum(!is.na(hh$momage)), "\n")

#-------------------------------
# factor coding for table1
#-------------------------------
hh <- hh %>%
  mutate(
    sex_child = factor(
      # use sexo_nino from survey
      sexo_nino,
      levels = c(1, 2),
      labels = c("Male", "Female")
    ),
    mom_ethnicity = factor(
      mom_ethnicity,
      levels = c(1, 2, 3, 4, 888),
      labels = c("Afro-Ecuadorian", "Chachi", "Mestizo", "Manabí", "Other")
    ),
    momedu_cat = factor(
      momedu_cat,
      levels = c("Primary or less", "Lower secondary",
                 "Upper secondary", "Post-secondary or greater")
    ),
    birthorder = factor(
      birthorder,
      levels = c("1", "2", "3+"),
      labels = c("First", "Second", "Third or later")
    ),
    hhcrowd = factor(
      hhcrowd,
      levels = c(0, 1),
      labels = c("< 3 persons/room", "≥ 3 persons/room")
    ),
    hhwealthquart = factor(
      hhwealthquart,
      levels = c(1, 2, 3, 4),
      labels = c("Q1 (lowest)", "Q2", "Q3", "Q4 (highest)")
    )
  )

label(hh$momage)        <- "Maternal age at childbirth (years)"
label(hh$mom_ethnicity) <- "Maternal ethnicity"
label(hh$momedu_cat)    <- "Maternal education"
label(hh$birthorder)    <- "Birth order"
label(hh$sex_child)     <- "Child sex"
label(hh$hhcrowd)       <- "Household crowding"
label(hh$hhwealthquart) <- "Wealth quartile"

t1a <- table1(
  ~ momage + mom_ethnicity + momedu_cat + birthorder +
    sex_child + hhcrowd + hhwealthquart | community_type_3,
  data = hh,
  overall = "Overall"
)
t1a

#-------------------------------
# save CSV version
#-------------------------------
t1a_df <- as.data.frame(t1a)
write_csv(t1a_df, here("output", "tables", "table1a_household_characteristics.csv"))
cat("Table 1a saved:", nrow(t1a_df), "rows\n")
```

---

# Table 1b: Cohort DBS and vaccination coverage summary

```{r table-1b}
#-------------------------------
# child-level summary (sex, community, DBS count, age at first DBS)
#-------------------------------
# add in sexo_nino from survey data
d0 <- d0 %>% 
  left_join(survey %>% select(house_ID, sexo_nino), by = "house_ID",  relationship = "many-to-one")

child_sum <- d0 %>%
  filter(pathogen == "Tetanus") %>%         # one row per child-visit
  group_by(childid) %>%
  summarise(
    community_type_3  = first(community_type_3),
    sex               = first(sexo_nino),
    n_dbs             = n_distinct(timepoint),
    age_first_dbs_mo  = min(agemonths_best, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    sex_lbl = factor(sex, levels = c(1, 2), labels = c("Male", "Female"))
  )

cat("Sex distribution:\n"); print(table(child_sum$sex_lbl))
cat("\nDBS per child — median:", median(child_sum$n_dbs),
    "IQR:", paste(quantile(child_sum$n_dbs, c(0.25, 0.75)), collapse = "–"), "\n")
cat("Age at first DBS (months) — median:", round(median(child_sum$age_first_dbs_mo), 1),
    "IQR:", paste(round(quantile(child_sum$age_first_dbs_mo, c(0.25, 0.75)), 1), collapse = "–"), "\n")

#-------------------------------
# vaccination coverage by timepoint and vaccine
# using visit-level indicators (dose received prior to DBS collection)
# denominator: children with non-missing visit-specific vaccination info
#-------------------------------
d_vax <- d0 %>%
  filter(pathogen == "Tetanus") %>%   # unique child-visit rows; vax vars same across pathogens
  mutate(
    penta_any_n = if_else(penta1_visit == "YES", 1L, 0L),   # ≥1 penta dose (tetanus/diphtheria)
    mmr_any_n   = if_else(mmr1_visit   == "YES", 1L, 0L)    # ≥1 MMR dose  (rubella/measles)
  )

cov_by_tp <- d_vax %>%
  group_by(timepoint) %>%
  summarise(
    penta_n_vacc  = sum(penta_any_n, na.rm = TRUE),
    penta_n_rec   = sum(!is.na(penta_any_n)),
    mmr_n_vacc    = sum(mmr_any_n, na.rm = TRUE),
    mmr_n_rec     = sum(!is.na(mmr_any_n)),
    .groups = "drop"
  ) %>%
  mutate(
    penta_pct = penta_n_vacc / penta_n_rec * 100,
    mmr_pct   = mmr_n_vacc   / mmr_n_rec   * 100,
    penta_cell = sprintf("%d/%d (%.1f%%)", penta_n_vacc, penta_n_rec, penta_pct),
    mmr_cell   = sprintf("%d/%d (%.1f%%)", mmr_n_vacc,   mmr_n_rec,   mmr_pct)
  )

#-------------------------------
# assemble Table 1b
#-------------------------------
t1b_rows <- tibble(
  Characteristic = c(
    "Children enrolled, n",
    "Sex, n (%)",
    "  Male",
    "  Female",
    "Community type, n (%)",
    "  Esmeraldas City",
    "  Borbon",
    "  Rural Villages",
    "DBS samples per child, median (IQR)",
    "Age at first DBS (months), median (IQR)",
    "Vaccination coverage at each age — ≥1 pentavalent dose (tetanus / diphtheria)",
    paste0("  Age ", cov_by_tp$timepoint, " months: ", cov_by_tp$penta_cell),
    "Vaccination coverage at each age — ≥1 MMR dose (rubella / measles)",
    paste0("  Age ", cov_by_tp$timepoint, " months: ", cov_by_tp$mmr_cell)
  )
)

n_total    <- n_distinct(child_sum$childid)
n_male     <- sum(child_sum$sex_lbl == "Male", na.rm = TRUE)
n_female   <- sum(child_sum$sex_lbl == "Female", na.rm = TRUE)
n_esm      <- sum(child_sum$community_type_3 == "Esmeraldas City")
n_bor      <- sum(child_sum$community_type_3 == "Borbon")
n_rur      <- sum(child_sum$community_type_3 == "Rural Villages")
dbs_med    <- median(child_sum$n_dbs)
dbs_iqr    <- quantile(child_sum$n_dbs, c(0.25, 0.75))
age_med    <- round(median(child_sum$age_first_dbs_mo), 1)
age_iqr    <- round(quantile(child_sum$age_first_dbs_mo, c(0.25, 0.75)), 1)

t1b <- tibble(
  Characteristic = c(
    "Children enrolled, n",
    "Sex, n (%)",
    sprintf("  Male, n (%%)"),
    sprintf("  Female, n (%%)"),
    "Community type, n (%)",
    "  Esmeraldas City",
    "  Borbon",
    "  Rural Villages",
    "DBS samples per child, median (IQR)",
    "Age at first DBS (months), median (IQR)"
  ),
  Value = c(
    as.character(n_total),
    "",
    sprintf("%d (%.1f%%)", n_male,   n_male / n_total * 100),
    sprintf("%d (%.1f%%)", n_female, n_female / n_total * 100),
    "",
    sprintf("%d (%.1f%%)", n_esm, n_esm / n_total * 100),
    sprintf("%d (%.1f%%)", n_bor, n_bor / n_total * 100),
    sprintf("%d (%.1f%%)", n_rur, n_rur / n_total * 100),
    sprintf("%d (%d–%d)",  dbs_med, dbs_iqr[1], dbs_iqr[2]),
    sprintf("%.1f (%.1f–%.1f)", age_med, age_iqr[1], age_iqr[2])
  )
)

# append vaccination coverage rows
cov_penta_rows <- tibble(
  Characteristic = c(
    "Vaccination coverage — ≥1 pentavalent dose received by visit (tetanus, diphtheria)^a^",
    paste0("  Age ", cov_by_tp$timepoint, " months")
  ),
  Value = c("", cov_by_tp$penta_cell)
)

cov_mmr_rows <- tibble(
  Characteristic = c(
    "Vaccination coverage — ≥1 MMR dose received by visit (rubella, measles)^a^",
    paste0("  Age ", cov_by_tp$timepoint, " months")
  ),
  Value = c("", cov_by_tp$mmr_cell)
)

t1b_full <- bind_rows(t1b, cov_penta_rows, cov_mmr_rows)

write_csv(t1b_full, here("output", "tables", "table1b_cohort_summary.csv"))
cat("Table 1b saved:", nrow(t1b_full), "rows\n")

t1b_full %>%
  gt() %>%
  tab_header(title = "Table 1b. Cohort Summary") %>%
  tab_footnote("^a^ Coverage = n with ≥1 dose received before the DBS collection visit / n attending with vaccination information available (%). Vaccination records may be missing for some visits.") %>%
  cols_label(Characteristic = "Characteristic", Value = "Overall")
```

---

# Tables 2–5: Seroprevalence by age and community type

```{r seroprev-tables, results='asis'}
for (pg in levels(d0$pathogen)) {

  sp_long <- seroprev_cells(d0, pg)

  # wide table: rows = timepoint, columns = community
  sp_wide <- sp_long %>%
    mutate(community_type_3 = as.character(community_type_3)) %>%
    pivot_wide_table("timepoint_lbl", "community_type_3", "cell", comm_col_order)

  # row order
  row_order <- c("6 months", "9 months", "12 months", "18 months", "24 months", "Overall")
  sp_wide <- sp_wide %>%
    mutate(timepoint_lbl = factor(timepoint_lbl, levels = row_order)) %>%
    arrange(timepoint_lbl)

  # save CSV (long format with all metrics)
  sp_csv <- sp_long %>%
    mutate(
      community_type_3 = as.character(community_type_3),
      pathogen_name = pg
    ) %>%
    select(pathogen_name, timepoint_lbl, community_type_3,
           n_pos, n_test, prev_pct, ci_lo, ci_hi, cell)

  tbl_num <- which(levels(d0$pathogen) == pg) + 1
  fname   <- sprintf("table%d_seroprev_%s.csv", tbl_num, tolower(pg))
  write_csv(sp_csv, here("output", "tables", fname))
  cat("Saved:", fname, "—", nrow(sp_csv), "rows\n")

  # display
  gt_tbl <- sp_wide %>%
    gt(rowname_col = "timepoint_lbl") %>%
    tab_header(title = sprintf("Table %d. %s IgG seroprevalence by age and community", tbl_num, pg)) %>%
    tab_stubhead(label = "Age group") %>%
    tab_footnote(
      "Cell: n seropositive / n tested (%) [95% exact binomial CI]. Overall row pools across all visit ages; overall column pools across all communities. Repeated-measures structure not accounted for in pooled rows/columns."
    )

  print(gt_tbl)
  cat("\n\n")
}
```

---

# Tables 6–7: Vaccination coverage by dose, age, and community type

```{r coverage-tables}
#-------------------------------
# vaccination dataset: one row per child-visit (filter to one pathogen)
# vaccination variables are identical across pathogens
#-------------------------------
d_cov <- d0 %>% filter(pathogen == "Tetanus")

#-------------------------------
# Table 6: Pentavalent vaccine (Penta 1, 2, 3)
#-------------------------------
penta_doses <- list(
  "Penta 1" = "penta1_visit",
  "Penta 2" = "penta2_visit",
  "Penta 3" = "penta3_visit"
)

t6_long <- map_dfr(names(penta_doses), function(dose_lbl) {
  coverage_cells(d_cov, penta_doses[[dose_lbl]], dose_lbl)
})

# row order
row_order <- c("6 months", "9 months", "12 months", "18 months", "24 months", "Overall")

t6_wide <- t6_long %>%
  mutate(community_type_3 = as.character(community_type_3)) %>%
  pivot_wide_table("timepoint_lbl", "community_type_3", "cell", comm_col_order) %>%
  mutate(
    timepoint_lbl = factor(timepoint_lbl, levels = row_order),
    dose = t6_long$dose[match(timepoint_lbl, t6_long$timepoint_lbl)]
  )

# re-derive dose column after pivot (pivot drops it)
t6_wide <- t6_long %>%
  mutate(community_type_3 = as.character(community_type_3)) %>%
  select(dose, timepoint_lbl, community_type_3, cell) %>%
  pivot_wider(names_from = community_type_3, values_from = cell) %>%
  select(dose, timepoint_lbl, any_of(comm_col_order)) %>%
  mutate(timepoint_lbl = factor(timepoint_lbl, levels = row_order)) %>%
  arrange(dose, timepoint_lbl)

# CSV (long)
t6_csv <- t6_long %>%
  mutate(community_type_3 = as.character(community_type_3)) %>%
  select(dose, timepoint_lbl, community_type_3, n_pos, n_test, prev_pct, ci_lo, ci_hi, cell)
write_csv(t6_csv, here("output", "tables", "table6_coverage_pentavalent.csv"))
cat("Table 6 saved:", nrow(t6_csv), "rows\n")

# display
t6_wide %>%
  gt(rowname_col = "timepoint_lbl", groupname_col = "dose") %>%
  tab_header(title = "Table 6. Pentavalent vaccine coverage by dose, age, and community") %>%
  tab_stubhead(label = "Age group") %>%
  tab_footnote(
    "Cell: n vaccinated / n with vaccination record (%) [95% exact binomial CI]. Vaccinated = dose received prior to DBS collection visit. Overall row pools across all visit ages."
  )
```

```{r table7}
#-------------------------------
# Table 7: MMR vaccine (MMR 1, MMR 2)
#-------------------------------
mmr_doses <- list(
  "MMR 1" = "mmr1_visit",
  "MMR 2" = "mmr2_visit"
)

t7_long <- map_dfr(names(mmr_doses), function(dose_lbl) {
  coverage_cells(d_cov, mmr_doses[[dose_lbl]], dose_lbl)
})

t7_wide <- t7_long %>%
  mutate(community_type_3 = as.character(community_type_3)) %>%
  select(dose, timepoint_lbl, community_type_3, cell) %>%
  pivot_wider(names_from = community_type_3, values_from = cell) %>%
  select(dose, timepoint_lbl, any_of(comm_col_order)) %>%
  mutate(timepoint_lbl = factor(timepoint_lbl, levels = row_order)) %>%
  arrange(dose, timepoint_lbl)

t7_csv <- t7_long %>%
  mutate(community_type_3 = as.character(community_type_3)) %>%
  select(dose, timepoint_lbl, community_type_3, n_pos, n_test, prev_pct, ci_lo, ci_hi, cell)
write_csv(t7_csv, here("output", "tables", "table7_coverage_mmr.csv"))
cat("Table 7 saved:", nrow(t7_csv), "rows\n")

t7_wide %>%
  gt(rowname_col = "timepoint_lbl", groupname_col = "dose") %>%
  tab_header(title = "Table 7. MMR vaccine coverage by dose, age, and community") %>%
  tab_stubhead(label = "Age group") %>%
  tab_footnote(
    "Cell: n vaccinated / n with vaccination record (%) [95% exact binomial CI]. Vaccinated = dose received prior to DBS collection visit. Overall row pools across all visit ages."
  )
```

---

# Figure 1: Seroprevalence and vaccination coverage by age (overall)

```{r fig1-data}
#-------------------------------
# Panel A data: seroprevalence by age (overall, all 4 pathogens)
#-------------------------------

# colors follow Stuart's draft: Tetanus=cbpal[4], Diphtheria=cbpal[2],
#   Rubella=cbpal[3], Measles=cbpal[8]
pathogen_colors <- c(
  "Tetanus"    = cbpal[7],   # red    — pentavalent-preventable (warm)
  "Diphtheria" = cbpal[2],   # orange — pentavalent-preventable (warm)
  "Rubella"    = cbpal[3],   # light blue — MMR-preventable (cool)
  "Measles"    = cbpal[6]    # dark blue  — MMR-preventable (cool)
)

sp_overall <- d0 %>%
  group_by(pathogen, timepoint) %>%
  summarise(
    n_pos  = sum(seropos, na.rm = TRUE),
    n_test = sum(!is.na(seropos)),
    .groups = "drop"
  ) %>%
  add_binom_ci()

#-------------------------------
# Panel B data: vaccination coverage by dose and age (all 5 doses)
#-------------------------------
dose_vars <- list(
  "Penta 1" = "penta1_visit",
  "Penta 2" = "penta2_visit",
  "Penta 3" = "penta3_visit",
  "MMR 1"   = "mmr1_visit",
  "MMR 2"   = "mmr2_visit"
)

cov_overall <- map_dfr(names(dose_vars), function(dose_lbl) {
  d_cov %>%
    mutate(vacc = if_else(.data[[dose_vars[[dose_lbl]]]] == "YES", 1L, 0L)) %>%
    group_by(timepoint) %>%
    summarise(
      n_pos  = sum(vacc, na.rm = TRUE),
      n_test = sum(!is.na(vacc)),
      .groups = "drop"
    ) %>%
    mutate(dose = dose_lbl)
}) %>%
  add_binom_ci()

# factor doses in anticipated coverage order at 24 months (Penta 1 highest, MMR 2 lowest)
cov_24 <- cov_overall %>%
  filter(timepoint == 24) %>%
  arrange(desc(prev_pct)) %>%
  pull(dose)
cov_overall <- cov_overall %>%
  mutate(dose = factor(dose, levels = cov_24))

dose_colors <- c(
  "Penta 1" = penta_pal[["1"]],
  "Penta 2" = penta_pal[["2"]],
  "Penta 3" = penta_pal[["3"]],
  "MMR 1"   = mmr_pal[["1"]],
  "MMR 2"   = mmr_pal[["2"]]
)
```

```{r fig1-plot, fig.width=8, fig.height=8}
pd <- position_dodge(width = 0.3)

panel_a <- sp_overall %>%
  ggplot(aes(x = timepoint, y = prev_pct, color = pathogen, group = pathogen)) +
  geom_linerange(aes(ymin = ci_lo, ymax = ci_hi), position = pd) +
  geom_point(size = 2, position = pd) +
  geom_line(position = pd) +
  scale_color_manual(values = pathogen_colors, name = NULL) +
  scale_x_continuous(breaks = seq(6, 24, by = 3)) +
  scale_y_continuous(breaks = seq(0, 100, 20), limits = c(0, 100)) +
  labs(x = "Age (months)", y = "Seroprevalence (%)") +
  theme_minimal() +
  theme(panel.grid.minor.x = element_blank(),
        legend.position = "bottom")

panel_b <- cov_overall %>%
  ggplot(aes(x = timepoint, y = prev_pct,
             color = dose, group = dose)) +
  geom_linerange(aes(ymin = ci_lo, ymax = ci_hi), position = pd) +
  geom_point(size = 2, position = pd) +
  geom_line(position = pd) +
  scale_color_manual(values = dose_colors, name = "Dose") +
  scale_x_continuous(breaks = seq(6, 24, by = 3)) +
  scale_y_continuous(breaks = seq(0, 100, 20), limits = c(0, 100)) +
  labs(x = "Age (months)", y = "Vaccination coverage (%)") +
  guides(color = guide_legend(nrow = 2, byrow = TRUE)) +   # Penta row / MMR row
  theme_minimal() +
  theme(panel.grid.minor.x = element_blank(),
        legend.position = "bottom")

# aspect.ratio = 1 makes each panel's plotting region square (≈1:1), matching
# the panel proportions in Figures 2 and 3
fig1 <- (panel_a / panel_b) +
  plot_annotation(tag_levels = "A") &
  theme(aspect.ratio = 1, plot.tag = element_text(face = "bold"))

fig1

ggsave(
  filename = here("output", "figures", "fig1_seroprev_coverage_byage.png"),
  plot = fig1,
  width = 130, height = 240, units = "mm", dpi = 300
)
cat("Figure 1 saved.\n")
```

---

# Figure 2: Seroprevalence by age and community type

```{r fig2-data}
sp_comm <- d0 %>%
  group_by(pathogen, community_type_3, timepoint) %>%
  summarise(
    n_pos  = sum(seropos, na.rm = TRUE),
    n_test = sum(!is.na(seropos)),
    .groups = "drop"
  ) %>%
  add_binom_ci()

# urban -> rural gradient: city = red, Borbon = blue, rural = green (forest).
# red and green are non-adjacent (blue between them) in the ordered legend.
comm_colors <- c(
  "Esmeraldas City" = cbpal[7],   # red
  "Borbon"          = cbpal[6],   # blue
  "Rural Villages"  = cbpal[4]    # green (rainforest)
)
# legend order: urban -> rural
comm_breaks <- c("Esmeraldas City", "Borbon", "Rural Villages")
```

```{r fig2-plot, fig.width=9, fig.height=8}
make_comm_panel <- function(data, pathogen_name) {
  data %>%
    filter(pathogen == pathogen_name) %>%
    ggplot(aes(x = timepoint, y = prev_pct,
               color = community_type_3, group = community_type_3)) +
    geom_linerange(aes(ymin = ci_lo, ymax = ci_hi)) +
    geom_point(size = 2) +
    geom_line() +
    scale_color_manual(values = comm_colors, breaks = comm_breaks,
                       name = NULL) +
    scale_x_continuous(breaks = seq(6, 24, by = 3)) +
    scale_y_continuous(breaks = seq(0, 100, 20), limits = c(0, 100)) +
    labs(x = "Age (months)", y = "Seroprevalence (%)",
         title = pathogen_name) +
    theme_minimal() +
    theme(panel.grid.minor.x = element_blank())
}

p_tet  <- make_comm_panel(sp_comm, "Tetanus")
p_diph <- make_comm_panel(sp_comm, "Diphtheria")
p_rub  <- make_comm_panel(sp_comm, "Rubella")
p_mea  <- make_comm_panel(sp_comm, "Measles")

fig2 <- (p_tet | p_diph) / (p_rub | p_mea) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "A") &
  theme(legend.position = "bottom", legend.direction = "horizontal",
        plot.tag = element_text(face = "bold"))

fig2

ggsave(
  filename = here("output", "figures", "fig2_seroprev_bycommunity.png"),
  plot = fig2,
  width = 200, height = 180, units = "mm", dpi = 300
)
cat("Figure 2 saved.\n")
```

---

# Figure 3: Vaccination coverage by dose and community type

```{r fig3-data}
cov_comm <- map_dfr(names(dose_vars), function(dose_lbl) {
  d_cov %>%
    mutate(vacc = if_else(.data[[dose_vars[[dose_lbl]]]] == "YES", 1L, 0L)) %>%
    group_by(community_type_3, timepoint) %>%
    summarise(
      n_pos  = sum(vacc, na.rm = TRUE),
      n_test = sum(!is.na(vacc)),
      .groups = "drop"
    ) %>%
    mutate(dose = dose_lbl)
}) %>%
  add_binom_ci()
```

```{r fig3-plot, fig.width=11, fig.height=8}
make_dose_panel <- function(data, dose_name) {
  data %>%
    filter(dose == dose_name) %>%
    ggplot(aes(x = timepoint, y = prev_pct,
               color = community_type_3, group = community_type_3)) +
    geom_linerange(aes(ymin = ci_lo, ymax = ci_hi)) +
    geom_point(size = 2) +
    geom_line() +
    scale_color_manual(values = comm_colors, breaks = comm_breaks,
                       name = NULL) +
    scale_x_continuous(breaks = seq(6, 24, by = 3)) +
    scale_y_continuous(breaks = seq(0, 100, 20), limits = c(0, 100)) +
    labs(x = "Age (months)", y = "Coverage (%)",
         title = sprintf("Pentavalent dose %s",
                         gsub("Penta ", "", dose_name))) +
    theme_minimal() +
    theme(panel.grid.minor.x = element_blank())
}

make_mmr_panel <- function(data, dose_name) {
  data %>%
    filter(dose == dose_name) %>%
    ggplot(aes(x = timepoint, y = prev_pct,
               color = community_type_3, group = community_type_3)) +
    geom_linerange(aes(ymin = ci_lo, ymax = ci_hi)) +
    geom_point(size = 2) +
    geom_line() +
    scale_color_manual(values = comm_colors, breaks = comm_breaks,
                       name = NULL) +
    scale_x_continuous(breaks = seq(6, 24, by = 3)) +
    scale_y_continuous(breaks = seq(0, 100, 20), limits = c(0, 100)) +
    labs(x = "Age (months)", y = "Coverage (%)",
         title = sprintf("MMR dose %s", gsub("MMR ", "", dose_name))) +
    theme_minimal() +
    theme(panel.grid.minor.x = element_blank())
}

p3_penta1 <- make_dose_panel(cov_comm, "Penta 1")
p3_penta2 <- make_dose_panel(cov_comm, "Penta 2")
p3_penta3 <- make_dose_panel(cov_comm, "Penta 3")
p3_mmr1   <- make_mmr_panel(cov_comm, "MMR 1")
p3_mmr2   <- make_mmr_panel(cov_comm, "MMR 2")

fig3 <- wrap_plots(
  p3_penta1, p3_penta2, p3_penta3,
  p3_mmr1,   p3_mmr2,   guide_area(),
  ncol = 3
) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "A") &
  theme(plot.tag = element_text(face = "bold"))

fig3

ggsave(
  filename = here("output", "figures", "fig3_coverage_bycommunity.png"),
  plot = fig3,
  width = 260, height = 175, units = "mm", dpi = 300
)
cat("Figure 3 saved.\n")
```

---

# Supplemental: pairwise community differences in seroprevalence by age

*Supplemental / held in reserve.* For each pathogen, we estimate the marginal
difference in seroprevalence between each pair of community types at each visit
age. Models are fit with `geeglm` (linear probability model: gaussian family /
identity link, so coefficients are differences in prevalence directly), with an
independence working correlation and cluster-robust SEs clustered by child —
consistent with the other inferential analyses. Because the community × age
model is saturated, the point estimates equal the observed cell-proportion
differences; the GEE supplies cluster-robust inference. Pairwise contrasts and
95% CIs are obtained with `emmeans` (Tukey-adjusted across the three community
pairs within each age).

```{r supp-community-diffs}
#-------------------------------
# short labels for the 3 pairwise contrasts
#-------------------------------
contrast_recode <- c(
  "Esmeraldas City - Borbon"         = "City − Borbon",
  "Esmeraldas City - Rural Villages" = "City − Rural",
  "Borbon - Rural Villages"          = "Borbon − Rural"
)

#-------------------------------
# per-pathogen LPM GEE + pairwise community contrasts at each age
#-------------------------------
supp_notes <- character(0)
diff_list <- list()
for (pg in levels(d0$pathogen)) {
  dat <- d0 %>%
    filter(pathogen == pg, !is.na(seropos)) %>%
    arrange(childid) %>%
    mutate(cid = as.integer(factor(childid)), seropos = as.integer(seropos))

  m <- withCallingHandlers(
    geeglm(seropos ~ community_type_3 * timept_f, family = gaussian,
           id = cid, corstr = "independence", data = dat),
    warning = function(w) {
      supp_notes[[length(supp_notes) + 1]] <<- sprintf("%s: %s", pg, conditionMessage(w))
      invokeRestart("muffleWarning")
    }
  )

  prs <- as.data.frame(
    summary(pairs(emmeans(m, ~ community_type_3 | timept_f)), infer = c(TRUE, TRUE))
  )

  diff_list[[pg]] <- prs %>%
    transmute(
      pathogen = pg,
      age      = as.character(timept_f),
      contrast = contrast_recode[as.character(contrast)],
      diff_pct = estimate * 100,
      ci_lo    = lower.CL * 100,
      ci_hi    = upper.CL * 100,
      p_value  = p.value
    )
}

diff_df <- bind_rows(diff_list) %>%
  mutate(
    pathogen     = factor(pathogen, levels = levels(d0$pathogen)),
    contrast_lbl = factor(contrast, levels = c("City − Borbon", "City − Rural", "Borbon − Rural")),
    age          = factor(age, levels = c("6", "9", "12", "18", "24")),
    # set CI to NA when diff is 0 to suppress error bars in plot (just one point)
    ci_lo        = ifelse(diff_pct == 0, NA, ci_lo),
    ci_hi        = ifelse(diff_pct == 0, NA, ci_hi)
  )

#-------------------------------
# save CSV (Table S1)
#-------------------------------
diff_csv <- diff_df %>%
  transmute(
    pathogen, age, contrast,
    `Difference (95% CI), pct points` = sprintf("%.1f (%.1f, %.1f)", diff_pct, ci_lo, ci_hi),
    p_value = if_else(p_value < 0.001, "<0.001", formatC(p_value, format = "g", digits = 3))
  )
write_csv(diff_csv, here("output", "tables", "tableS1_seroprev_diff_by_community.csv"))
cat("Table S1 saved:", nrow(diff_csv), "rows (4 pathogens x 5 ages x 3 contrasts)\n")

if (length(supp_notes) > 0) {
  cat("GEE notes:\n"); for (n in supp_notes) cat(" -", n, "\n")
} else {
  cat("All supplemental GEE models fit without warnings.\n")
}
```

```{r supp-forest, fig.width=9, fig.height=8}
#-------------------------------
# forest plot of pairwise differences, one panel per pathogen
# (x = age; y = difference in prevalence; reference line at 0; CI crossing 0 = n.s.)
#-------------------------------
contrast_cols <- c("City − Borbon" = cbpal[2],
                   "City − Rural"  = cbpal[3],
                   "Borbon − Rural" = cbpal[8])

# common y-axis range + 10-point breaks, harmonized across all four panels
y_rng    <- range(c(diff_df$ci_lo, diff_df$ci_hi), na.rm = TRUE)
y_lim    <- c(floor(y_rng[1] / 10) * 10, ceiling(y_rng[2] / 10) * 10)
y_breaks <- seq(y_lim[1], y_lim[2], by = 10)

# ylab supplied only for the left-column panels (A, C); B and D share the y-axis
make_forest <- function(pg, ylab = NULL) {
  diff_df %>%
    filter(pathogen == pg) %>%
    ggplot(aes(x = age, y = diff_pct, color = contrast_lbl)) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
    geom_pointrange(aes(ymin = ci_lo, ymax = ci_hi),
                    position = position_dodge(width = 0.6), size = 0.3) +
    scale_color_manual(values = contrast_cols, name = NULL) +
    scale_y_continuous(breaks = y_breaks) +
    coord_cartesian(ylim = y_lim) +
    labs(x = "Age (months)", y = ylab, title = pg) +
    theme_minimal() +
    theme(panel.grid.minor = element_blank())
}

figS1 <- (make_forest("Tetanus", "Difference in seroprevalence (%)") | make_forest("Diphtheria")) /
         (make_forest("Rubella", "Difference in seroprevalence (%)") | make_forest("Measles")) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "A") &
  theme(legend.position = "bottom", plot.tag = element_text(face = "bold"))

figS1

ggsave(
  filename = here("output", "figures", "figS1_seroprev_diff_by_community.png"),
  plot = figS1,
  width = 220, height = 185, units = "mm", dpi = 300
)
cat("Figure S1 saved.\n")
```

---

# Session information

```{r session-info}
sessionInfo()
```

