---
title: "Concordance Between Vaccination Records and Serology"
author: "Stuart Torres, Ben Arnold"
date: "Updated `r Sys.Date()`"
format:
  html:
    toc: true
    embed-resources: true
    code-fold: true
    link-external-newwindow: true
---

# Preamble

```{r preamble, message=FALSE}
library(here)
source(here("R", "0-config.R"))
```

# Load and prepare data

```{r load-data}
#-------------------------------
# read processed dataset
#-------------------------------
d0 <- read_rds(here("data", "ecomid_sero_vpd_2026-06-11.rds"))

#-------------------------------
# pathogen factor; collapse community type to 3 levels (for Table 9 strata)
#-------------------------------
d0 <- d0 %>%
  mutate(
    pathogen = factor(pathogen,
                      levels = c("Tetanus", "Diphtheria", "Rubella", "Measles")),
    community_type_3 = factor(
      case_when(
        community_name == "Esmeraldas" ~ "Esmeraldas City",
        community_name == "Borbon"     ~ "Borbon",
        TRUE                           ~ "Rural Villages"
      ),
      levels = c("Esmeraldas City", "Borbon", "Rural Villages"))
  )

#-------------------------------
# define index-test "vaccinated" status per pathogen:
#   received >= 1 dose of the relevant vaccine prior to DBS collection
#   pentavalent (tetanus, diphtheria); MMR (rubella, measles)
#-------------------------------
d0 <- d0 %>%
  mutate(
    vac = case_when(
      pathogen %in% c("Tetanus", "Diphtheria") ~ as.integer(penta_doses >= 1),
      pathogen %in% c("Rubella", "Measles")    ~ as.integer(mmr_doses >= 1)
    )
  )

cat("Vaccinated (>=1 dose) defined from penta_doses / mmr_doses.\n")
```

```{r gee-helpers}
#-------------------------------
# collect GEE convergence / estimability notes for the report
#-------------------------------
gee_notes <- character(0)

#-------------------------------
# helper: fit a GEE logistic model with an INDEPENDENCE working correlation
# and cluster-robust (sandwich) SEs clustered by child (geeglm default).
# data must contain an integer cluster id column `cid`, sorted by child.
#-------------------------------
fit_gee <- function(formula, data, tag) {
  withCallingHandlers(
    tryCatch(
      geeglm(formula, family = binomial(link = "logit"),
             id = cid, corstr = "independence", data = data),
      error = function(e) {
        gee_notes[[length(gee_notes) + 1]] <<-
          sprintf("%s: GEE error: %s", tag, conditionMessage(e))
        NULL
      }
    ),
    warning = function(w) {
      gee_notes[[length(gee_notes) + 1]] <<-
        sprintf("%s: GEE warning: %s", tag, conditionMessage(w))
      invokeRestart("muffleWarning")
    }
  )
}

#-------------------------------
# helper: predicted probability + 95% cluster-robust CI for a linear
# combination L of the GEE coefficients (Wald CI on the logit scale).
#-------------------------------
gee_pred_ci <- function(fit, L) {
  b   <- coef(fit)
  V   <- vcov(fit)                       # robust sandwich covariance (geeglm default)
  eta <- as.numeric(L %*% b)
  se  <- sqrt(as.numeric(t(L) %*% V %*% L))
  c(est = plogis(eta), lo = plogis(eta - 1.96 * se), hi = plogis(eta + 1.96 * se))
}

#-------------------------------
# helper: format a metric as "73.4% (60.1, 84.2)"
# (the raw counts are already shown in the 2x2 columns, so no [n/N] here)
#-------------------------------
fmt_metric <- function(v) {
  if (is.na(v["est"])) return("—")
  sprintf("%.1f%% (%.1f, %.1f)",
          v["est"] * 100, v["lo"] * 100, v["hi"] * 100)
}

#-------------------------------
# helper: full concordance summary for one pathogen subset.
#   serology (seropos) = gold standard; vaccination record (vac) = index test
#
#   sensitivity, specificity  <- model  vaccinated ~ seropos
#   PPV, NPV                   <- model  seropos ~ vaccinated
#   (specificity and NPV are the complements of the respective fitted cells)
#-------------------------------
concordance_row <- function(dat, label) {
  dat <- dat %>%
    filter(!is.na(seropos), !is.na(vac)) %>%
    arrange(childid) %>%
    mutate(cid = as.integer(factor(childid)),
           seropos = as.integer(seropos),
           vac = as.integer(vac))

  TP <- sum(dat$vac == 1 & dat$seropos == 1)
  FP <- sum(dat$vac == 1 & dat$seropos == 0)
  FN <- sum(dat$vac == 0 & dat$seropos == 1)
  TN <- sum(dat$vac == 0 & dat$seropos == 0)

  na_vec <- c(est = NA, lo = NA, hi = NA)

  # Model A: vaccinated ~ seropos  -> sensitivity, specificity
  # specificity = complement of P(vaccinated | seronegative): flip CI bounds.
  # as.numeric() strips inner names so c(est=, lo=, hi=) stays clean.
  fitA <- fit_gee(vac ~ seropos, dat, paste(label, "sens/spec"))
  if (!is.null(fitA)) {
    sens <- gee_pred_ci(fitA, c(1, 1))                 # P(vac=1 | sero=1)
    s0   <- gee_pred_ci(fitA, c(1, 0))                 # P(vac=1 | sero=0)
    spec <- c(est = 1 - as.numeric(s0["est"]),
              lo  = 1 - as.numeric(s0["hi"]),
              hi  = 1 - as.numeric(s0["lo"]))
  } else {
    sens <- na_vec; spec <- na_vec
  }

  # Model B: seropos ~ vaccinated -> PPV, NPV
  # NPV = complement of P(seropositive | not vaccinated): flip CI bounds.
  fitB <- fit_gee(seropos ~ vac, dat, paste(label, "PPV/NPV"))
  if (!is.null(fitB)) {
    ppv <- gee_pred_ci(fitB, c(1, 1))                  # P(sero=1 | vac=1)
    n0  <- gee_pred_ci(fitB, c(1, 0))                  # P(sero=1 | vac=0)
    npv <- c(est = 1 - as.numeric(n0["est"]),
             lo  = 1 - as.numeric(n0["hi"]),
             hi  = 1 - as.numeric(n0["lo"]))
  } else {
    ppv <- na_vec; npv <- na_vec
  }

  # Overall agreement: P(record agrees with serology) = (TP + TN) / N,
  # from an intercept-only GEE on the agreement indicator (same independence +
  # cluster-robust SE approach as the other metrics).
  N_tot <- TP + FP + FN + TN
  datG  <- dat %>% mutate(agree = as.integer(vac == seropos))
  fitG  <- fit_gee(agree ~ 1, datG, paste(label, "agreement"))
  agr   <- if (!is.null(fitG)) gee_pred_ci(fitG, 1) else na_vec

  tibble(
    Group = label,
    N = N_tot,
    `TN (IgG-/Vac-)` = TN,
    `FN (IgG+/Vac-)` = FN,
    `FP (IgG-/Vac+)` = FP,
    `TP (IgG+/Vac+)` = TP,
    Agreement = fmt_metric(agr),
    Sensitivity = fmt_metric(sens),
    Specificity = fmt_metric(spec),
    PPV = fmt_metric(ppv),
    NPV = fmt_metric(npv),
    small_cell = any(c(TP, FP, FN, TN) < 5)
  )
}
```

---

# Table 8: Concordance across all four pathogens

Serology is the gold standard; the vaccination record (≥ 1 dose received prior to
the visit) is the index test. Tetanus and diphtheria use all visit ages; rubella
and measles are restricted to the 12-, 18-, and 24-month visits (MMR not given
earlier under Ecuador's national schedule). 95% CIs are from GEE logistic models
with an independence working correlation and cluster-robust SEs clustered by child.

```{r table8}
#-------------------------------
# track excluded observations (missing serology or vaccination status)
#-------------------------------
excl_summary <- tibble()
t8_rows <- list()

for (pg in levels(d0$pathogen)) {
  dp <- d0 %>% filter(pathogen == pg)
  if (pg %in% c("Rubella", "Measles")) {
    dp <- dp %>% filter(timepoint %in% c(12, 18, 24))
  }

  n_before <- nrow(dp)
  n_excl   <- sum(is.na(dp$seropos) | is.na(dp$vac))
  excl_summary <- bind_rows(excl_summary,
    tibble(Pathogen = pg, n_eligible = n_before,
           n_excluded = n_excl, n_analyzed = n_before - n_excl))

  t8_rows[[pg]] <- concordance_row(dp, pg) %>% rename(Pathogen = Group)
}
t8 <- bind_rows(t8_rows)

cat("Observations excluded for missing serology or vaccination status:\n")
print(as.data.frame(excl_summary), row.names = FALSE)
cat("\nTotal excluded across pathogens:", sum(excl_summary$n_excluded), "\n")

#-------------------------------
# small-cell flag, save CSV, display
#-------------------------------
t8_display <- t8 %>%
  mutate(Pathogen = if_else(small_cell, paste0(Pathogen, " †"), Pathogen)) %>%
  select(-small_cell)

write_csv(t8_display, here("output", "tables", "table8_concordance.csv"))
cat("\nTable 8 saved.\n")

t8_display %>%
  gt() %>%
  tab_header(
    title = "Table 8. Concordance of vaccination record with serology",
    subtitle = "Serology = gold standard; vaccination record (≥1 dose) = index test"
  ) %>%
  tab_spanner(label = "2×2 counts",
              columns = c(`TN (IgG-/Vac-)`, `FN (IgG+/Vac-)`,
                          `FP (IgG-/Vac+)`, `TP (IgG+/Vac+)`)) %>%
  tab_spanner(label = "Diagnostic accuracy (% [n/N], 95% CI)",
              columns = c(Agreement, Sensitivity, Specificity, PPV, NPV)) %>%
  tab_footnote("% is the GEE point estimate (equals the crude proportion under the independence working correlation); 95% CIs are cluster-robust (sandwich), clustered by child. Sensitivity/specificity from model `vaccinated ~ seropos`; PPV/NPV from model `seropos ~ vaccinated`. Numerators/denominators for each metric are given by the 2×2 counts.") %>%
  tab_footnote("Agreement = overall percent agreement between vaccination record and serology = (TP + TN) / N, from an intercept-only GEE on the agreement indicator.") %>%
  tab_footnote("Rubella and measles restricted to 12-, 18-, and 24-month visits.") %>%
  tab_footnote("† one or more 2×2 cells < 5; interpret with caution.")
```

---

# Table 9: Measles vs. rubella seropositivity (12–24 months)

Measles and rubella antigens are co-delivered in the MMR vaccine. Lower measles
seropositivity than rubella among the same children could signal a problem with
the heat-labile measles component (e.g., cold-chain failure). The comparison is
shown overall and stratified by community type, at the 12-, 18-, and 24-month
visits.

```{r table9}
#-------------------------------
# data for the comparison: rubella + measles at 12/18/24m, non-missing seropos
#-------------------------------
dcmp <- d0 %>%
  filter(pathogen %in% c("Rubella", "Measles"),
         timepoint %in% c(12, 18, 24)) %>%
  mutate(pathogen = factor(pathogen, levels = c("Rubella", "Measles")))

n_excl_cmp <- sum(is.na(dcmp$seropos))
dcmp <- dcmp %>% filter(!is.na(seropos)) %>% mutate(seropos = as.integer(seropos))

cat("Child-visits (per pathogen) analyzed:", nrow(dcmp) / 2, "\n")
cat("Unique children:", n_distinct(dcmp$childid), "\n")
cat("Excluded for missing serology:", n_excl_cmp, "\n")

#-------------------------------
# helper: measles vs rubella comparison for one (sub)set —
# per-pathogen seroprevalence (exact binomial CI) + GEE risk difference
# (rubella - measles), independence + cluster-robust SE, marginalized via emmeans
#-------------------------------
mr_compare <- function(dat, label) {
  dat <- dat %>% arrange(childid) %>% mutate(cid = as.integer(factor(childid)))

  st <- dat %>%
    group_by(pathogen) %>%
    summarise(n_pos = sum(seropos), n_test = n(), .groups = "drop") %>%
    rowwise() %>%
    mutate(pct_ci = sprintf("%.1f%% (%.1f%% to %.1f%%)",
                            n_pos / n_test * 100,
                            binom.test(n_pos, n_test)$conf.int[1] * 100,
                            binom.test(n_pos, n_test)$conf.int[2] * 100)) %>%
    ungroup()

  rd_str <- tryCatch({
    fit <- geeglm(seropos ~ pathogen, family = binomial(link = "logit"),
                  id = cid, corstr = "independence", data = dat)
    rd  <- summary(pairs(regrid(emmeans(fit, ~ pathogen))), infer = c(TRUE, TRUE))
    pf  <- if (rd$p.value < 0.001) "< 0.001" else paste("=", formatC(rd$p.value, format = "g", digits = 3))
    sprintf("%.1f%% (%.1f, %.1f); p %s",
            rd$estimate * 100, rd$lower.CL * 100, rd$upper.CL * 100, pf)
  }, error = function(e) "not estimable")

  tibble(
    Stratum = label,
    `Rubella N`          = st$n_test[st$pathogen == "Rubella"],
    `Rubella % (95% CI)` = st$pct_ci[st$pathogen == "Rubella"],
    `Measles N`          = st$n_test[st$pathogen == "Measles"],
    `Measles % (95% CI)` = st$pct_ci[st$pathogen == "Measles"],
    `Difference (Rubella − Measles)` = rd_str
  )
}

#-------------------------------
# overall + stratified by community type
#-------------------------------
t9 <- bind_rows(
  mr_compare(dcmp, "Overall"),
  mr_compare(dcmp %>% filter(community_type_3 == "Esmeraldas City"), "Esmeraldas City"),
  mr_compare(dcmp %>% filter(community_type_3 == "Borbon"),          "Borbon"),
  mr_compare(dcmp %>% filter(community_type_3 == "Rural Villages"),  "Rural Villages")
)

write_csv(t9, here("output", "tables", "table9_measles_rubella_comparison.csv"))
cat("Table 9 saved.\n")

t9 %>%
  gt() %>%
  tab_header(
    title = "Table 9. Measles vs. rubella seropositivity (12–24 months)",
    subtitle = "Same children, both antigens co-delivered in MMR"
  ) %>%
  tab_spanner(label = "Rubella", columns = c(`Rubella N`, `Rubella % (95% CI)`)) %>%
  tab_spanner(label = "Measles", columns = c(`Measles N`, `Measles % (95% CI)`)) %>%
  cols_label(
    `Rubella N` = "N", `Rubella % (95% CI)` = "Seropositive % (95% CI)",
    `Measles N` = "N", `Measles % (95% CI)` = "Seropositive % (95% CI)"
  ) %>%
  tab_footnote("N = number of child-visits tested; Seropositive % with 95% exact binomial CI. Difference: risk difference (rubella − measles) from a GEE logistic model (independence working correlation, cluster-robust SEs by child), marginalized to the probability scale; 95% CI and p-value shown.") %>%
  tab_footnote("Measles and rubella are measured on the same DBS at each visit (paired; within-child-visit seropositivity correlation ≈ 0.73). Clustering on child means the cluster-robust SE already accounts for this pairing — it matches a paired (McNemar) analysis — so no separate paired test is needed.")
```

---

# GEE model notes

```{r gee-notes}
if (length(gee_notes) == 0) {
  cat("All GEE models converged with no warnings or estimability issues.\n")
} else {
  cat("GEE model notes (", length(gee_notes), "):\n", sep = "")
  for (nt in gee_notes) cat(" -", nt, "\n")
}
```

---

# Session information

```{r session-info}
# print and also save session info (reporting convention, spec §6)
writeLines(capture.output(sessionInfo()), here("output", "session_info.txt"))
sessionInfo()
```
